\Large \bf Cyclone: Warp Speed Replication for Key Value Stores

### 2017-04-20 paper/paper.tex
Updates to paper.

In order to recover from a crash, we make two important assumptions about the
underlying SSD. First, we assume that 4KB is the \emph{minimum} atomic unit for
updating pages on the SSD even under power failure i.e. there are no shorn
writes on a 4KB page~\cite{shorn_writes}. We also assume that the SSD has power
loss data protection meaning that writes cached in the drive's volatile cache
are written to the SSD using a backup capacitor in the event of power failure -
a property of many data center class SSDs today, including the Intel DC P3600
SSD we use in our evaluation. Together these two assumptions mean that we can
recover a consistent prefix of the log on a power failure.

@@

The fact that we can immedately absorb updates into an NVM log and only move
them to the flashlog in suitable units for block IO solves a common problem with
logging to secondary storage - the need to do group commit. Logging systems
often wait to collect a batch of entries to write to secondary storage trading
latency for throughput. This is unnecessary in Cyclone as it
effectively makes use of directly attached NVM with a two level log that removes
the need to make this tradeoff.

\section{Horizontal Scaling}

\subsection{API Commutativity}

@@

>>>>

In order to recover from a crash, we make two important assumptions about the
underlying SSD. First, we assume that 4KB is the \emph{minimum} atomic unit for
updating pages on the SSD even under power failure i.e. there are no shorn
writes on a 4KB page~\cite{shorn_writes}. We also assume that the SSD has power
loss data protection meaning that writes cached in the drive's volatile cache
are written to the SSD using a backup capacitor in the event of power failure -
a property of many data center class SSDs today, including the Intel DC P3600
SSD~\cite{ssd_spec} we use in our evaluation. Together these two assumptions
mean that we can recover a consistent prefix of the log on a power failure.

@@

The fact that we can immedately absorb updates into an NVM log and only move
them to the flashlog in suitable units for block IO, sidesteps a common problem
with logging to secondary storage - the need to do group commit. Logging systems
often wait to collect a batch of entries to write to secondary storage trading
latency for throughput. This is unnecessary in Cyclone as it effectively makes
use of directly attached NVM with a two level log that removes the need to make
this tradeoff.

\begin{figure}
  \centering
  \includegraphics[scale=0.3]{figures2/race.pdf}
  \caption{Race with multiple physical logs}
  \label{fig:race}
\end{figure}

\section{Horizontal Scaling}
The basic principle in Cyclone is to turn log replication into a software packet
switching problem. A key ingredient in software packet switching is applying
multiple CPU cores to parallelize the control plane, scaling message processing
throughput. In this section we describe how we can horizontally scale Cyclone's
single \emph{logical} log across multiple \emph{physical} logs. For each
physical log, we run an independent version of RAFT with the corresponding
control plane mapped to a dedicated core. Further, as we shortly describe we
assume a concurrent key value store where a set of CPU cores independently run
operations from the physical logs on the shared memory key-value store data
structure, synchronizing as necessary.

We begin by describing how a key-value store API can be mapped on these
independent physical logs without divergence in replica state. We then discuss
how different pieces of work related to replication and the key value store
itself are distributed across available CPU cores and available NIC queues.

\subsection{Leveraging API Commutativity}
Update requests to a key-value store consist of a key and an associated operation -
such as ``put'' that set the value for the key or ``merge'' that calls a
user-defined operator to merge a new value for the key into the existing value.
A naive approach that maps a key to a random physical log and subsequently to a
random core on the system for execution runs into the determinism problem -
replicas could diverge with respect to the final value assigned to a key on
different replicas. We illustrate this problem with an example in
Figure~\ref{fig:race} where two updates to the same key race with each other on
different physical logs and execution cores to end up being applied in different
orders on different replicas, leading to a divergence of state.

The solution to this problem is observe that the key-value store is a concurrent
data structure that ultimately admits a serialization of operations to
it. Further, consecutive operations to different keys in the serial history can
be interchanged, because operations to different keys are
\emph{commutative}. The correct way to use different physical logs is therefore
to simply hash the key to select the physical log \emph{and} the core on which
the operation executes. This guarantees determinism as the serial history of
two different replicas can be transformed into the same serial history by
reordering operations to different keys.

Further relaxations to the mapping scheme are possible when operations to the
same key are also commutative. A good example is when key value stores are used
to maintain counters. Different increment operations to the same counter are
commutative if one is prepared to tolerate non-determinism in the sequence of
updates seen by a reader with the counter reaching the same eventual value. This
specific use-case, in fact, lead to the addition of the merge operation in
Rocksdb, a key value store that we use for evaluation. Merge operations
therefore could be dispatched to a random physical log with no harm. For this
paper however, we stick on the stricter requirement that operations to the same
key are not commutative.

@@

TBD: mention no linearizability for reads at the moment
